Back to ``Normal'' for the Disintegrating Planet Candidate KIC 12557548 b

### 2018-07-06 ms.tex
edits to abstract and intro.

\begin{abstract}
KIC 12557548 b is one of several disintegrating planet candidates which lose mass in the form of dust particles escaping their surface.
Photometric monitoring of KIC 12557548 b originally showed evidence for a slowdown of disintegration activity in 2013 and 2014.
Here, we report 2016 measurements that are consistent with photometry from the Kepler spacecraft in 2009-2013.
We also evaluate high resolution archival spectra from the Subaru HDS spectrograph and find spectra consistent with a main-sequence T=4500 $\pm$110 K star.
This disfavors the hypothesis that planet disintegration affected previous high resolution spectra results.
We apply PCA to the Kepler long cadence data to understand the modes of disintegration.
There is a tentative 491 day periodicity of the second principal component, which corresponds to possible long-term evolution of the dust grain sizes, though the mechanism is unclear.
\end{abstract}

@@

\section{Introduction}
\citet{rappaport} first reported the discovery of a highly unusual stellar system in the Kepler field.
KIC 12557548's (hereafter \shStar's) light curve exhibits flux dips with a period of 0.653 days.
Surprisingly, the flux dips are neither symmetric in shape nor constant in amplitude, varying from $\sim$ 0\% to $\sim$ 1.2\%.
The amplitudes of the flux dips are highly stochastic and unpredictable from one event to the next 0.653 days later.
This indicates the creation and clearing of material that can cover up to 1\% of the star in $\lesssim$ 1 day timescales.
These unusual light curve behaviors are best explained as dust extinction from material that is escaping a rocky planet (KIC 12557548 b, hereafter \sha) in a short period orbit.
The KIC 12557548 system (hereafter \shStar) has the exciting possibility of being an opportunity to study a planet that has been peeled away layer by layer to reveal an interior core.

The discovery of K2-22b \citep{sanchis-ojedak2-22}, WD 1145+017 \citep{vanderburg2015wdDisintegrating}, KOI 2700 \citep{rappaport2014KOI2700} showed that other systems have similar light curves and that other disintegrating planet systems may be useful laboratories for understanding planet evolution and composition.
Multi-wavelength light curve measurements show that the dust particles can have wavelength-dependent transmission, as predicted from sub-micron dust particles \citep{bochinski2015evolving,sanchis-ojedak2-22}.

Furthermore, there are perplexing systems that also display stochastic transit behavior but for which there are still no fully agreed-upon explanations: KIC 8462852 (ie. Boyajian's star) \citep{boyajian846}, RIK-210/EPIC 205483258 \citep{david2017rik210} and PTFO 8-8695 \citep{vanEyken2012ptfTTauri}.
KIC 8462852 may be a family of comets from one or more parent bodies that recently were broken apart \citep{boyajian846,bodman2016kic8462852cometFamily,wright2016familiesOfSolutions}.
RIK-210 and PTFO 8-8695 are both in young systems ($\lesssim$ 10 and $\lesssim$3 Myr respectively) so they may be proto-planet candidates or structures within a protoplanetary disk \citep{stauffer2017orbitingClouds,yu2015ptfO8d8695PlanetHypothesisTest}.
RIK-210 could be accreting material that causes variable extinction while PTFO 8-8695's orbit may be precessing and causing variations in transit depth depending on its longitude, though only portions of the lightcurve are explained this way \citep{barnes2013PTFO8d8695NodalPrecession,yu2015ptfO8d8695PlanetHypothesisTest}.

The underlying planet candidate \sha\ has not been detected directly, but several observational constraints have placed upper limits on its size and mass.
High precision radial velocity measurements of the host star with Keck/HIRES put an upper limit on the reflex motion due to the planet and constrains the mass to be $\lesssim 1.2 M_{\mathrm Jup}$ \citep{croll2014}.
\citet{masuda2018rvKIC1255} find that the upper limit on the radial velocity semi-amplitude is 86 m/s corresponding to a planet mass $\lesssim 0.28 M_{\mathrm Jup}$ using the Subaru High Dispersion Spectrometer (HDS).
There is no detection of secondary eclipses of \sha\ with a 3$\sigma$ upper limit of $5 \times 10^{-5}$ in the Kepler bandpass, which means that for an albedo of 0.5, the radius of the planet must be smaller than 4600 km \citep{vanWerkhoven2014}.
A radiative-hydrodynamic wind model predicts that the mass loss rate will be a strong function of planet mass and for the present-day mass loss rate of $\sim$ M$_\oplus/Gyr$, the modeled mass is 0.014 $M_\oplus$.

@@

\subsection{Principal Component Analysis}
\begin{figure*}[!hbtp]
\begin{centering}
\includegraphics[width=0.32\textwidth]{images/kepler/photometry_riverplot.pdf}
\includegraphics[width=0.32\textwidth]{images/kepler/pca_model_2D.pdf}
\includegraphics[width=0.32\textwidth]{images/kepler/pca_resid_2D.pdf}
\caption{Cleaned Kepler Long-cadence data (Left) and model fit using 4 Principal Components (Middle).
The residuals (Right) have $\lesssim 0.1\%$ residuals in contrast to the average-light curve method used in Figure \ref{fig:riverPlots}.
}\label{fig:riverPlotsPCA}
\end{centering}
\end{figure*}

@@

We obtained ground-based R band photometry from the Kuiper 61-inch observatory to follow up a possible slowdown in disintegration activity observed in 2013 and 2014.
\citet{schlawin2016kic1255} found shallow transits (all below the Kepler observatory average) in August-September 2013 and August-September 2014.
One possibility was a long term evolution of the disintegration activity, such as a reduction of available dusty material of the planet or a slowdown in the disintegration mechanism.
Our transit depths from June-July 2016 are instead consistent with the depths from the Kepler observatory from 2009 to 2013, with a KS-test between the two distributions having a p-value of 0.5 that depths are drawn from the same distribution.
This indicates that the shallow transit depths observed in August-September 2013 and August-September 2014 likely fell into the 14-36 day long quiescent intervals of disintegration activity as observed in Kepler photometry \citep{kawahara2013starspots,vanWerkhoven2014,croll2015starspots}.

>>>>

\begin{abstract}
KIC 12557548 b is first of a growing class of intriguing disintegrating planet candidates, which lose mass in the form of dust particles escaping their surface.
Here, we follow up two perplexing observations of the system: 1) the transits appeared shallower than average in 2013 and 2014 and 2) the parameters derived from a high resolution spectrum of the star differed from other results using photometry and low resolution spectroscopy.
We observe 5 transits of the system with the 61-inch Kuiper observatory in 2016 and show that they are consistent with photometry from the Kepler spacecraft in 2009-2013.
We also evaluate high resolution archival spectra from the Subaru HDS spectrograph and find spectra consistent with a main-sequence T=4500 $\pm$110 K star.
This disfavors the hypothesis that planet disintegration affected previous high resolution spectra results.
We apply PCA to the Kepler long cadence data to understand the modes of disintegration.
There is a tentative 491 day periodicity of the second principal component, which corresponds to possible long-term evolution of the dust grain sizes, though the mechanism on such long timescales remains unclear.
\end{abstract}

@@

\section{Introduction}
\citet{rappaport} first reported the discovery of a highly unusual stellar system in the Kepler field.
KIC 12557548's (hereafter \shStar's) light curve exhibits flux dips with a period of 0.653 days.
Surprisingly, the flux dips are neither symmetric in shape nor constant in amplitude, varying from $\sim$ 0\% to $\sim$ 1.2\%.
The amplitudes of the flux dips are highly stochastic and unpredictable from one event to the next 0.653 days later.
This indicates that dusty material with a projected area $\sim$1\% of the stellar surface is created and cleared in $\lesssim$ 1 day timescales.
These unusual light curve behaviors are best explained as dust extinction from material that is escaping a rocky planet (KIC 12557548 b, hereafter \sha) in a short period orbit.
The KIC 12557548 system (hereafter \shStar) has the exciting possibility of being an opportunity to study a planet that has been peeled away layer by layer to reveal an interior core.

The discovery of K2-22b \citep{sanchis-ojedak2-22}, WD 1145+017 \citep{vanderburg2015wdDisintegrating}, KOI 2700 \citep{rappaport2014KOI2700} and KIC 3542116 \citep{rappaport2018exocomets} showed that other systems have similar light curves and that several disintegrating planet systems may be useful laboratories for understanding planet evolution and composition.
Multi-wavelength light curve measurements show that the dust particles can have wavelength-dependent transmission, as predicted from sub-micron dust particles \citep{bochinski2015evolving,sanchis-ojedak2-22}.

Furthermore, there are perplexing systems that also display stochastic transit behavior but for which there are still no fully agreed-upon explanations: KIC 8462852 (ie. Boyajian's star) \citep{boyajian846}, RIK-210/EPIC 205483258 \citep{david2017rik210},  and PTFO 8-8695 \citep{vanEyken2012ptfTTauri}.
KIC 8462852 may be a family of comets from one or more parent bodies that recently were broken apart \citep{boyajian846,bodman2016kic8462852cometFamily,wright2016familiesOfSolutions}.
RIK-210 and PTFO 8-8695 are both in young systems ($\lesssim$ 10 and $\lesssim$3 Myr respectively) so they may be proto-planet candidates or structures within a protoplanetary disk \citep{stauffer2017orbitingClouds,yu2015ptfO8d8695PlanetHypothesisTest}.
RIK-210 could be accreting material that causes variable extinction while PTFO 8-8695's orbit may be precessing and causing variations in transit depth depending on its longitude, though only portions of the lightcurve are explained this way \citep{barnes2013PTFO8d8695NodalPrecession,yu2015ptfO8d8695PlanetHypothesisTest}.

The underlying planet candidate \sha\ has not been detected directly, but several observational constraints have placed upper limits on its size and mass.
High precision radial velocity measurements of the host star with Keck/HIRES put an upper limit on the reflex motion due to the planet and constrains the mass to be $\lesssim 1.2 M_{\mathrm Jup}$ \citep{croll2014}.
\citet{masuda2018rvKIC1255} find that the upper limit on the radial velocity semi-amplitude is 86 m/s corresponding to a planet mass $\lesssim 0.28 M_{\mathrm Jup}$ using the Subaru High Dispersion Spectrometer (HDS).
There is no detection of secondary eclipses of \sha\ with a 3$\sigma$ upper limit of $5 \times 10^{-5}$ in the Kepler bandpass, which means that for an albedo of 0.5, the radius of the planet must be smaller than 4600 km \citep{vanWerkhoven2014}.
A radiative-hydrodynamic wind model predicts that the mass loss rate will be a strong function of planet mass and for the present-day mass loss rate of $\sim$ M$_\oplus/Gyr$, the modeled mass is 0.014 $M_\oplus$ \citep{perez-becker}.

@@

\subsection{Principal Component Analysis}
\begin{figure*}[!hbtp]
\begin{centering}
\includegraphics[width=0.32\textwidth]{images/kepler/photometry_riverplot.pdf}
\includegraphics[width=0.32\textwidth]{images/kepler/pca_model_2D.pdf}
\includegraphics[width=0.32\textwidth]{images/kepler/pca_resid_2D.pdf}
\caption{Cleaned Kepler Long-cadence data (Left) and model fit using 4 Principal Components (Middle).
The residuals (Right) are $\lesssim 0.1\%$, significantly smaller than the $\sim 0.2\%$ residuals when using the average-light curve method shown in Figure \ref{fig:riverPlots}.
}\label{fig:riverPlotsPCA}
\end{centering}
\end{figure*}

@@

We obtained ground-based R band photometry from the Kuiper 61-inch observatory to follow up a possible slowdown in disintegration activity observed in 2013 and 2014.
\citet{schlawin2016kic1255} found shallow transits that were all below the Kepler observatory average in August-September 2013 and August-September 2014.
One possibility was a long term evolution of the disintegration activity, such as a reduction of available dusty material of the planet or a slowdown in the disintegration mechanism.
Our transit depths from June-July 2016 are instead consistent with the depths from the Kepler observatory from 2009 to 2013, with a KS-test between the two distributions having a p-value of 0.5 that depths are drawn from the same distribution.
This indicates that the shallow transit depths observed in August-September 2013 and August-September 2014 likely fell into the 14-36 day long quiescent intervals of disintegration activity as observed in Kepler photometry \citep{kawahara2013starspots,vanWerkhoven2014,croll2015starspots}.
